---
title: "Sealice microbiome"
author: "Bela Klimesova"
date: "2025-12-12"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

# Loading microbiome datasets

```{r}
library(phyloseq)
library(ggplot2)
suppressMessages(library(vegan))
suppressMessages(library(microbiomeMarker))
library(agricolae)
suppressMessages(library(tidyverse))
library(patchwork)
suppressMessages(library(ggvenn))
```

Data loading: Microbiome read count data

```{r}
Sequen<-read.table("Sealice_Microbiome_Reads.csv",
	header = TRUE,sep = ',',fill = TRUE,dec = ".",na.strings = "NA")
```

Adjustment of dataset and devision to Illumina and Nanopore data

```{r}
rownames(Sequen)<-Sequen[,1]
Sequen[,1]<- NULL

Sequen_t<-t(Sequen)

Illumina<-Sequen[,c(1:20)]
Nanopore<-Sequen[,c(21:40)]
```

Data loading: Microbiome metadata

```{r}
sample<-read.csv("Sealice_Microbiome_Metadata.txt",header=T,row.names=1,sep = "\t")
```

Adjustment of dataset and devision to Illumina and Nanopore data

```{r}
sample$time<-rep(c(rep('February',5),rep('March',5),rep('April',5),rep('May',5)),2)
sample$platform<-c(rep('Illumina',20),rep('Nanopore',20))

Ill_sample<-sample[sample$platform=='Illumina',]
Nan_sample<-sample[sample$platform=='Nanopore',]
```

Data loading: Microbiome taxonomy data

```{r}
observation<-read.csv("Sealice_Microbiome_Taxonomy.tsv",header=T,row.names=1, sep = "\t")
```

# Dataset format adjustment

Discarding global singletons in all three datasets (Illumina, Nanopore, Overall)

```{r}
Singletons<-names(apply(Illumina,1,sum)[apply(Illumina,1,sum)<2])
Illumina_1<-Illumina[!(rownames(Illumina)%in% Singletons),]

Singletons<-names(apply(Nanopore,1,sum)[apply(Nanopore,1,sum)<2])
Nanopore_1<-Nanopore[!(rownames(Nanopore)%in% Singletons),]

Singletons<-names(apply(Sequen,1,sum)[apply(Sequen,1,sum)<2])
Sequen_1<-Sequen[!(rownames(Sequen)%in% Singletons),]
```

Data to phyloseq format

Illumina
```{r}
OTU1 = otu_table(as.matrix(t(Illumina_1)), taxa_are_rows = FALSE) 
SAM1 = sample_data(Ill_sample, errorIfNULL = TRUE) 
TAX1 = tax_table(as.matrix(observation))
Ill_phylo <- phyloseq(OTU1, TAX1, SAM1)
```
Nanopore
```{r}
OTU2 = otu_table(as.matrix(t(Nanopore_1)), taxa_are_rows = FALSE) 
SAM2 = sample_data(Nan_sample, errorIfNULL = TRUE) 
TAX2 = tax_table(as.matrix(observation))
Nan_phylo <- phyloseq(OTU2, TAX2, SAM2)
```
Overall
```{r}
OTU = otu_table(as.matrix(t(Sequen)), taxa_are_rows = FALSE) 
SAM = sample_data(sample, errorIfNULL = TRUE) 
TAX = tax_table(as.matrix(observation))
phylo <- phyloseq(OTU, TAX, SAM)
```

# Data analysis

# Rarefaction curves

Adjustment of data fortmat
```{r}
Illumina_t <- t(Illumina_1)
Nanopore_t <- t(Nanopore_1)
```

Names of samples and matching colors
```{r}
leg.txt <- c(paste0(1:5,'_Feb'),paste0(1:5,'_Mar'),
		paste0(1:5,'_Apr'),paste0(1:5,'_May'))
col_vector <- c("#661100","#7F261D","#993B3B","#B25059","#CC6677",
		    "#DDCC77","#AAB666","#77A155","#438C44","#117733",
		    "#88CCEE","#72A1D4","#5D77BB","#484CA1","#332288",
		    "#44AA99","#55A194","#669990","#77908C","#888888")
```

Figure of rarefaction curves
```{r fig1, fig.cap="Rarefaction curve"}
layout(matrix(c(1,2,3,3), nrow = 2, byrow = TRUE),
       heights = c(3, 1))  # top row bigger than bottom

par(mar = c(5, 4, 4, 2), cex = 1, las = 1)

# Illumina
rarecurve(Illumina_t, 
          step = 100, lwd = 2, cex = 0.8, 
          xlab = "Number of reads", 
          ylab = "Number of OTU observed",
          ylim = c(1, 350), xlim = c(1, 250000),
          col = col_vector, size = 1.5, 
          label = FALSE, main = "A) Illumina")

# Nanopore
rarecurve(Nanopore_t, 
          step = 100, lwd = 2, cex = 0.8, 
          xlab = "Number of reads", 
          ylab = "Number of OTU observed",
          ylim = c(1, 350), xlim = c(1, 250000),
          col = col_vector, size = 1.5, 
          label = FALSE, main = "B) Nanopore")

par(mar = c(0, 0, 0, 0))
plot.new()
legend("center",
       legend = leg.txt, lwd = 2, col = col_vector,
       box.lwd = 0.6, cex = 1.0, ncol = 5) 
```

Determination of sequencing depth

Illumina
```{r}
sum_seq <- colSums(Illumina)
range(sum_seq)
sum(sum_seq)
mean(sum_seq)
sd(sum_seq)
```

Nanopore
```{r}
sum_seq <- colSums(Nanopore)
range(sum_seq)
sum(sum_seq)
mean(sum_seq)
sd(sum_seq)
```
# Alpha diversity and richness

Calculation of alpha diversity

Illumina
```{r}
data_richness <- estimateR(Illumina_t)
data_evenness <- diversity(Illumina_t) / log(specnumber(Illumina_t))
data_shannon <- diversity(Illumina_t, index = "shannon")
data_simpson <- diversity(Illumina_t, index = "simpson")
data_alphadiv1 <- cbind(Ill_sample, t(data_richness), data_shannon, 
                        data_simpson, data_evenness)
```

Nanopore
```{r}
data_richness <- estimateR(Nanopore_t)
data_evenness <- diversity(Nanopore_t) / log(specnumber(Nanopore_t))
data_shannon <- diversity(Nanopore_t, index = "shannon")
data_simpson <- diversity(Nanopore_t, index = "simpson")
data_alphadiv2 <- cbind(Nan_sample, t(data_richness), data_shannon,
                        data_simpson, data_evenness)
```

Combining dataset
```{r}
data_alphadiv<-rbind(as.data.frame(data_alphadiv1),as.data.frame(data_alphadiv2))
data_alphadiv$time<-factor(data_alphadiv$time,levels=c('February','March','April','May'))
```

## Shannon index

Test normal distribution
```{r}
shapiro.test(data_alphadiv1$data_shannon)
shapiro.test(data_alphadiv2$data_shannon)
```

ANOVA test to assess differences
```{r}
aov_test1<-aov(data_shannon ~ time, data = data_alphadiv1)
summary(aov_test1)

aov_test2<-aov(data_shannon ~ time, data = data_alphadiv2)
summary(aov_test2)
```

Post-hoc Tukey test
```{r}
hsd_test1 <- TukeyHSD(aov_test1)  
hsd_res1 <- HSD.test(aov_test1, "time", group=T)$groups  
hsd_res1 

hsd_test2 <- TukeyHSD(aov_test2)  
hsd_res2 <- HSD.test(aov_test2, "time", group=T)$groups  
hsd_res2 
```

Preparation of data for plotting
```{r}
summary_data <- data_alphadiv %>%
  group_by(time, platform) %>%
  summarise(mean = mean(data_shannon, na.rm = TRUE),
            sd = sd(data_shannon, na.rm = TRUE),
            .groups = 'drop')

label_data <- summary_data %>%
  mutate(label = c("a", "ab", "a", "b", "b", "c", "ab", "bc"),  
         y = mean + sd + 0.15) 
```

Figure of Shannon index
```{r}
P1 <- ggplot(data_alphadiv, aes(x = time, y = data_shannon, fill = time,alpha=0.5)) +
	geom_errorbar(data = summary_data,
                aes(x = time, ymin = mean - sd, ymax = mean + sd, group = platform),
                position = position_dodge(width = 0.75),
                width = 0.2,
                color = "black",
                inherit.aes = FALSE)+
	geom_text(data = label_data,
            aes(x = time, y = y, label = label,group = platform),
            position = position_dodge(width = 0.75),
            inherit.aes = FALSE,
            size = 4) +
	geom_boxplot(position = position_dodge(width = 0.75), coef = 0,outlier.shape = NA) +
	scale_fill_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
	scale_color_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
	labs(title = 'A)  Shannon index', x= ' ', y= '') +
	theme_minimal()+
	facet_wrap(~platform)+
	theme(legend.position="none",
	axis.text.x = element_text(angle = 90, hjust = 1))
```

## Simpson index

Test normal distribution
```{r}
shapiro.test(data_alphadiv1$data_simpson)
shapiro.test(data_alphadiv2$data_simpson)
```

ANOVA test to assess differences
```{r}
aov_test1<-aov(data_simpson ~ time, data = data_alphadiv1)
summary(aov_test1)

aov_test2<-aov(data_simpson ~ time, data = data_alphadiv2)
summary(aov_test2)
```

Post-hoc Tukey test
```{r}
hsd_test1 <- TukeyHSD(aov_test1)  
hsd_res1 <- HSD.test(aov_test1, "time", group=T)$groups  
hsd_res1 

hsd_test2 <- TukeyHSD(aov_test2) 
hsd_res2 <- HSD.test(aov_test2, "time", group=T)$groups  
hsd_res2 
```

Preparation of data for plotting
```{r}
summary_data <- data_alphadiv %>%
  group_by(time, platform) %>%
  summarise(mean = mean(data_simpson, na.rm = TRUE),
            sd = sd(data_simpson, na.rm = TRUE),
            .groups = 'drop')

label_data <- summary_data %>%
  mutate(label = c("ab", "ab", "a", "a", "b", "b", "b", "b"),  
         y = mean + sd + 0.05) 
```

Figure of Simpson index
```{r}
P2 <- ggplot(data_alphadiv, aes(x = time, y = data_simpson, fill = time,alpha=0.5)) +
  geom_errorbar(data = summary_data,
                aes(x = time, ymin = mean - sd, ymax = mean + sd, group = platform),
                position = position_dodge(width = 0.75),
                width = 0.2,
                color = "black",
                inherit.aes = FALSE)+
  geom_text(data = label_data,
            aes(x = time, y = y, label = label,group = platform),
            position = position_dodge(width = 0.75),
            inherit.aes = FALSE,
            size = 4) +
  geom_boxplot(position = position_dodge(width = 0.75), coef = 0,outlier.shape = NA) +
	scale_fill_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
	scale_color_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
  labs(title = 'B)  Simpson index', x= ' ', y= '') +
  theme_minimal()+
	facet_wrap(~platform)+
	theme(legend.position="none",
	axis.text.x = element_text(angle = 90, hjust = 1))
```

## Observed richness

Test normal distribution
```{r}
shapiro.test(data_alphadiv1$S.obs)
shapiro.test(data_alphadiv2$S.obs)
```

ANOVA and Kruskal test to assess differences
```{r}
Krus<-kruskal.test(S.obs ~ time, data = data_alphadiv1)
pairwise.wilcox.test(data_alphadiv1$S.obs,data_alphadiv1$time,
                 p.adjust.method = "BH")
aov_test2<-aov(S.obs ~ time, data = data_alphadiv2)
summary(aov_test2)
```

Preparation of data for plotting
```{r}
summary_data <- data_alphadiv %>%
  group_by(time, platform) %>%
  summarise(mean = mean(S.obs, na.rm = TRUE),
            sd = sd(S.obs, na.rm = TRUE),
            .groups = 'drop')
```

Figure of observed richness
```{r}
P3 <- ggplot(data_alphadiv, aes(x = time, y = S.obs, fill = time,alpha=0.5)) +
  geom_errorbar(data = summary_data,
                aes(x = time, ymin = mean - sd, ymax = mean + sd, group = platform),
                position = position_dodge(width = 0.75),
                width = 0.2,
                color = "black",
                inherit.aes = FALSE) +
  geom_boxplot(position = position_dodge(width = 0.75), coef = 0,outlier.shape = NA) +
	scale_fill_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
	scale_color_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
  labs(title = 'C)   Observed richness', x = '', y = '') +
  theme_minimal()+
	facet_wrap(~platform)+
	theme(legend.position="none",
	axis.text.x = element_text(angle = 90, hjust = 1))
```

## ACE

Test normal distribution
```{r}
shapiro.test(data_alphadiv1$S.ACE)
shapiro.test(data_alphadiv2$S.ACE)
```

ANOVA test to assess differences
```{r}
aov_test1<-aov(S.ACE ~ time, data = data_alphadiv1)
summary(aov_test1)

aov_test2<-aov(S.ACE ~ time, data = data_alphadiv2)
summary(aov_test2)
```

Post-hoc Tukey test
```{r}
hsd_test1 <- TukeyHSD(aov_test1) 
hsd_res1 <- HSD.test(aov_test1, "time", group=T)$groups  
hsd_res1 

hsd_test2 <- TukeyHSD(aov_test2) 
hsd_res2 <- HSD.test(aov_test2, "time", group=T)$groups  
hsd_res2
```

Preparation of data for plotting
```{r}
summary_data <- data_alphadiv %>%
  group_by(time, platform) %>%
  summarise(mean = mean(S.ACE, na.rm = TRUE),
            sd = sd(S.ACE, na.rm = TRUE),
            .groups = 'drop')

label_data <- summary_data %>%
  mutate(label = c("ab", "", "ab", "", "a", "", "b", ""),  
         y = mean + sd + 40) 
```

Figure of ACE
```{r}
P4 <- ggplot(data_alphadiv, aes(x = time, y = S.ACE, fill = time,alpha=0.5)) +
  geom_errorbar(data = summary_data,
                aes(x = time, ymin = mean - sd, ymax = mean + sd, group = platform),
                position = position_dodge(width = 0.75),
                width = 0.2,
                color = "black",
                inherit.aes = FALSE) +
  geom_text(data = label_data,
            aes(x = time, y = y, label = label,group = time),
            position = position_dodge(width = 0.75),
            inherit.aes = FALSE,
            size = 4) +
  geom_boxplot(position = position_dodge(width = 0.75), coef = 0,outlier.shape = NA) +
	scale_fill_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
	scale_color_manual(values = c("#CC6677","#DDCC77","#117733","#332288")) +
  labs(title = 'D)   ACE', x = '', y = '') +
  theme_minimal()+
  facet_wrap(~platform)+
	theme(legend.position="none",
	axis.text.x = element_text(angle = 90, hjust = 1))
```

Combined plot of richness and diversity
```{r fig2, fig.cap="Diversity and richness"}
((P1 | P2)/(P3 | P4)) 
```

# Data normalization = calculation of abundance
```{r}
Ill_phylo_norm <- transform_sample_counts(Ill_phylo, function(ASV) ASV/sum(ASV) *100 )
Nan_phylo_norm <- transform_sample_counts(Nan_phylo, function(ASV) ASV/sum(ASV) *100 )
phylo_norm <- transform_sample_counts(phylo, function(ASV) ASV/sum(ASV) *100 )
```

# Data filtering
Keep only taxa (columns) where at least one sample has ≥ 0.03 abundance

```{r}
threshold <- 0.03
```

Illumina
```{r}
Ill_atu_filt <- otu_table(Ill_phylo_norm)[, colSums(otu_table(Ill_phylo_norm) 
                                                    >= threshold) > 0]
ncol(Ill_atu_filt)
ncol(otu_table(Ill_phylo_norm))
```

Nanopore
```{r}
Nan_atu_filt <- otu_table(Nan_phylo_norm)[, colSums(otu_table(Nan_phylo_norm) 
                                                    >= threshold) > 0]
ncol(Nan_atu_filt)
ncol(otu_table(Nan_phylo_norm))
```

Overall
```{r}
atu_filt <- otu_table(phylo_norm)[, colSums(otu_table(phylo_norm) 
                                            >= threshold) > 0]
ncol(atu_filt)
ncol(otu_table(phylo_norm))
```

Overview of detected taxonomical groups
Illumina
```{r}
Names<-colnames(Ill_atu_filt)
Selec<-data.frame(TAX1[rownames(TAX1) %in% Names,])

length(unique(Selec$Phylum))
length(unique(Selec$Class))
length(unique(Selec$Order))
```

Nanopore
```{r}
Names<-colnames(Nan_atu_filt)
Selec<-data.frame(TAX2[rownames(TAX2) %in% Names,])

length(unique(Selec$Phylum))
length(unique(Selec$Class))
length(unique(Selec$Order))
```

# Beta diversity

```{r}
set.seed(1782)
```

Changing the format of data
```{r}
SAM1$grouping<-c(rep('Feb and Mar',10),rep('Apr and May',10))
Ill_phylo_filt_1 <- phyloseq(otu_table(Ill_atu_filt), TAX1, SAM1)

SAM2$grouping<-c(rep('Feb and Mar',10),rep('Apr and May',10))
Nan_phylo_filt_1 <- phyloseq(otu_table(Nan_atu_filt), TAX2, SAM2)
phylo_filt_1 <- phyloseq(otu_table(atu_filt), TAX, SAM)
```

Principal coordinate analysis

Illumina
```{r}
pcoa_bc1 = ordinate(Ill_phylo_filt_1, "PCoA", "bray")
P1<-plot_ordination(Ill_phylo_filt_1, pcoa_bc1, color = "time") + 
	geom_point(size = 3,alpha=0.5) +
	theme_minimal()+
	scale_color_manual(values=c("#CC6677","#DDCC77","#117733","#332288"),
		breaks = c('February','March','April','May'))+
	labs(title = 'A)   Illumina', x= 'Component [40.6%]',
		y= 'Component [20.8%]') +
	theme(text = element_text(size = 13))+
  stat_ellipse(aes(group = grouping), linetype = 2,alpha=0.5)
```

Nanopore
```{r}
pcoa_bc2 = ordinate(Nan_phylo_filt_1, "PCoA", "bray") 
P2<-plot_ordination(Nan_phylo_filt_1, pcoa_bc2, color = "time") + 
	geom_point(size = 3,alpha=0.5) +
	theme_minimal()+
	scale_color_manual(values=c("#CC6677","#DDCC77","#117733","#332288"),
		breaks = c('February','March','April','May'))+
	labs(title = 'B)   Nanopore', x= 'Component [51.3%]',
		y= 'Component [22.7%]')+
	theme(text = element_text(size = 13)) +
  stat_ellipse(aes(group = grouping), linetype = 2,alpha=0.5)
```

Combining of figures for beta diversity
```{r fig3, fig.cap="Beta-diversity: PCoA"}
((P1 | P2 ))  +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

Testing differences between months with PERMANOVA
Illumina
```{r}
adonis2(Ill_atu_filt~time,data=Ill_sample, permutations=9999, method="bray")
```

Nanopore
```{r}
adonis2(Nan_atu_filt~time,data=Nan_sample, permutations=9999, method="bray")
```

# Microbiome composition: phylum level

Adjusting format of data
```{r}
observation$SampleID<-rownames(observation)
```

Illumina
```{r}
Ill_atu_filt_rar<-data.frame(t(Ill_atu_filt))
Ill_atu_filt_rar$SampleID<-rownames(Ill_atu_filt_rar)
```

Nanopore
```{r}
Nan_atu_filt_rar<-data.frame(t(Nan_atu_filt))
Nan_atu_filt_rar$SampleID<-rownames(Nan_atu_filt_rar)
```

Overall
```{r}
atu_filt_rar<-data.frame(t(atu_filt))
atu_filt_rar$SampleID<-rownames(atu_filt_rar)
```

Merging datasets
```{r}
Tax_Data1<-merge(Ill_atu_filt_rar,observation,all.x=TRUE,by='SampleID')
colnames(Tax_Data1)[2:21]<-paste0(rep(c(1:5),4),'_',sample$time[1:20])

Tax_Data2<-merge(Nan_atu_filt_rar,observation,all.x=TRUE,by='SampleID')
colnames(Tax_Data2)[2:21]<-paste0(rep(c(1:5),4),'_',sample$time[1:20])

Tax_Data<-merge(atu_filt_rar,observation,all.x=TRUE,by='SampleID')
```

Sum per Phyla
```{r}
Tax_Phylum1<- aggregate(Tax_Data1[,2:21],by=list(Phylum=Tax_Data1$Phylum), sum)
Tax_Phylum1$Phylum<-substr(Tax_Phylum1$Phylum,4,40)

Tax_Phylum2<- aggregate(Tax_Data2[,2:21],by=list(Phylum=Tax_Data2$Phylum), sum)
Tax_Phylum2$Phylum<-substr(Tax_Phylum2$Phylum,4,40)
```

Tidying datasets
Illumina
```{r}
Tax_Phylum1<-data.frame(t(Tax_Phylum1))
colnames(Tax_Phylum1)<-Tax_Phylum1[1,]
Tax_Phylum1<-Tax_Phylum1[2:21,]
Tax_Phylum1$month<-substr(rownames(Tax_Phylum1),3,10)
Tax_Phylum1$Sample_ID<-rownames(Tax_Phylum1)
Tax_Phylum1$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Phyll<- gather(Tax_Phylum1,Phylums,Abundance,
                   Actinomycetota:Verrucomicrobiota, factor_key=TRUE)
Tax_Phyll$Abundance_r<-round(as.numeric(Tax_Phyll$Abundance),2)
```

Nanopore
```{r}
Tax_Phylum2<-data.frame(t(Tax_Phylum2))
colnames(Tax_Phylum2)<-Tax_Phylum2[1,]
Tax_Phylum2<-Tax_Phylum2[2:21,]
Tax_Phylum2$month<-substr(rownames(Tax_Phylum2),3,10)
Tax_Phylum2$Sample_ID<-rownames(Tax_Phylum2)
Tax_Phylum2$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Phyl2<- gather(Tax_Phylum2,Phylums,Abundance,
                   Bacillota:Verrucomicrobiota, factor_key=TRUE)
Tax_Phyl2$Abundance_r<-round(as.numeric(Tax_Phyl2$Abundance),2)
```

Filtering: phyla with abundance lower than 0.5% described as others
```{r}
Tax_Phyll$Phylum<-as.vector(Tax_Phyll$Phylums)
Tax_Phyll[Tax_Phyll$Abundance_r<0.5,]$Phylum<-'Others'

Tax_Phyl2$Phylum<-as.vector(Tax_Phyl2$Phylums)
Tax_Phyl2[Tax_Phyl2$Abundance_r<0.5,]$Phylum<-'Others'
```

Combining datasets
```{r}
Tax_Phyll$platform<-'Illumina'
Tax_Phyl2$platform<-'Nanopore'
Tax_Phyl<-rbind(Tax_Phyll,Tax_Phyl2)
```

Adjustment of final dataset
```{r}
Tax_Phyl_N<-aggregate(Tax_Phyl$Abundance_r,list(month=Tax_Phyl$month,
			Phylum=Tax_Phyl$Phylum,
			platform=Tax_Phyl$platform),mean)
Tax_Phyl_N$month<-factor(Tax_Phyl_N$month,levels = c('February','March','April','May'))

sort(tapply(Tax_Phyl_N$x,Tax_Phyl_N$Phylum,sum))
Tax_Phyl_N$Phylum<-factor(Tax_Phyl_N$Phylum,
	levels = c("Myxococcota",'Others','Bacillota',"Thermodesulfobacteriota",
	"Bdellovibrionota","Cyanobacteriota",'Patescibacteria',"Campylobacterota",
	"Verrucomicrobiota","Bacteroidota","Pseudomonadota"))
```

Figure of bacterial compositon at phylum level
```{r}
P_Phyl<-ggplot(Tax_Phyl_N,aes(x = month, y = x, fill = Phylum)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = c("#888888","#44AA99","#882255","#6699CC",
	"#999933","#AA4499","#88CCEE","#332288","#117733","#CC6677","#DDCC77")) +
  labs(title= 'A)', x= 'Month', 
	y= 'Proportion of the total abundance')+
	facet_wrap(~platform) +
  theme_minimal()+ 
	theme(text = element_text(size = 17),
	axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r fig4, fig.cap="Microbiome composition: phylum level"}
(P_Phyl)
```

Range of abundance values
Illumina
```{r}
tapply(Tax_Phyll$Abundance_r,Tax_Phyll$Phylum,mean)
range(Tax_Phyll[Tax_Phyll$Phylum=='Pseudomonadota',]$Abundance_r)
range(Tax_Phyll[Tax_Phyll$Phylum=='Bacteroidota',]$Abundance_r)
```

Nanopore
```{r}
tapply(Tax_Phyl2$Abundance_r,Tax_Phyl2$Phylum,mean)
range(Tax_Phyl2[Tax_Phyl2$Phylum=='Pseudomonadota',]$Abundance_r)
range(Tax_Phyl2[Tax_Phyl2$Phylum=='Bacteroidota',]$Abundance_r)
```

# Bacterial composition: class level

Adjusting the dataset to class level
Illumina
```{r}
Tax_Data1[Tax_Data1$Class=="c__Incertae_Sedis",]$Class<-'p__NB1_j_Inc_Sed'

Tax_Class1<- aggregate(Tax_Data1[,2:21],by=list(Class=Tax_Data1$Class), sum)
Tax_Class1$Class<-substr(Tax_Class1$Class,4,40)

Tax_Class1<-data.frame(t(Tax_Class1))
colnames(Tax_Class1)<-Tax_Class1[1,]
Tax_Class1<-Tax_Class1[2:21,]
Tax_Class1$month<-substr(rownames(Tax_Class1),3,10)
Tax_Class1$Sample_ID<-rownames(Tax_Class1)
Tax_Class1$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Cla1<- gather(Tax_Class1,Class,
                  Abundance,Acidimicrobiia:NB1_j_Inc_Sed, factor_key=TRUE)
Tax_Cla1$Abundance_r<-round(as.numeric(Tax_Cla1$Abundance),4)
```

Nanopore
```{r}
Tax_Data2[Tax_Data2$Class=="c__Incertae_Sedis",]$Class<-'p__Marinimicrobia_inc._sed.'

Tax_Class2<- aggregate(Tax_Data2[,2:21],by=list(Class=Tax_Data2$Class), sum)
Tax_Class2$Class<-substr(Tax_Class2$Class,4,40)

Tax_Class2<-data.frame(t(Tax_Class2))
colnames(Tax_Class2)<-Tax_Class2[1,]
Tax_Class2<-Tax_Class2[2:21,]
Tax_Class2$month<-substr(rownames(Tax_Class2),3,10)
Tax_Class2$Sample_ID<-rownames(Tax_Class2)
Tax_Class2$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Cla2<- gather(Tax_Class2,Class,Abundance,
                  Alphaproteobacteria:Marinimicrobia_inc._sed., factor_key=TRUE)
Tax_Cla2$Abundance_r<-round(as.numeric(Tax_Cla2$Abundance),4)
```

Filtering only 10 orders with the highest abundance
```{r}
Part<-names(sort(tapply(as.numeric(Tax_Cla1$Abundance),Tax_Cla1$Class,mean)))[4:13]
Tax_Cla1$Class_N<-as.character(Tax_Cla1$Class)
Tax_Cla1[!(Tax_Cla1$Class_N %in% Part),]$Class_N<-'Others'
Tax_Cla1$platform<-'Illumina'

Part<-names(sort(tapply(as.numeric(Tax_Cla2$Abundance),Tax_Cla2$Class,mean)))[8:17]
Tax_Cla2$Class_N<-as.character(Tax_Cla2$Class)
Tax_Cla2[!(Tax_Cla2$Class_N %in% Part),]$Class_N<-'Others'
Tax_Cla2$platform<-'Nanopore'
```

Combining datasets
```{r}
Tax_Cla<-rbind(Tax_Cla1,Tax_Cla2)

Tax_Cla_New<-aggregate(Tax_Cla$Abundance_r,list(month=Tax_Cla$month,
			Class=Tax_Cla$Class_N,platform=Tax_Cla$platform),mean)

Tax_Cla_New$month<-factor(Tax_Cla_New$month,levels=c('February','March','April','May'))
Tax_Cla_New$Class<-factor(Tax_Cla_New$Class,
	levels = c('Others','Desulfuromonadia','Polyangiia','Bacteriovoracia',
		'Cyanobacteriia','Gracilibacteria','Campylobacteria',
		'Verrucomicrobiia','Alphaproteobacteria','Bacteroidia',
		'Gammaproteobacteria'))
sort(tapply(Tax_Cla_New$x,Tax_Cla_New$Class,mean))
```

Figure of bacterial compositon at class level
```{r}
P_Cla<-ggplot(Tax_Cla_New,aes(x = month, y = x, fill = Class)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = c("#332288","#117733","#CC6677","#DDCC77","#888888",
		"#6699CC","#882255","#999933","#44AA99","#AA4499","#88CCEE"
		)) +
  labs(title= 'C)', x= 'Month')+
	ylab('')+
  theme_minimal()+ 
	theme(axis.text.x = element_text(vjust = 0.5, hjust=1))+
	facet_wrap(~platform)+ 
	theme(text = element_text(size = 17),
	axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r fig5, fig.cap="Microbiome composition: class level"}
P_Cla
```

Range of abundance values
Illumina
```{r}
tapply(Tax_Cla1$Abundance_r,Tax_Cla1$Class,mean)
range(Tax_Cla1[Tax_Cla1$Class=='Gammaproteobacteria',]$Abundance_r)
range(Tax_Cla1[Tax_Cla1$Class=='Bacteroidia',]$Abundance_r)
range(Tax_Cla1[Tax_Cla1$Class=='Alphaproteobacteria',]$Abundance_r)
```

Nanopore
```{r}
tapply(Tax_Cla2$Abundance_r,Tax_Cla2$Class,mean)
range(Tax_Cla2[Tax_Cla2$Class=='Gammaproteobacteria',]$Abundance_r)
range(Tax_Cla2[Tax_Cla2$Class=='Bacteroidia',]$Abundance_r)
range(Tax_Cla2[Tax_Cla2$Class=='Alphaproteobacteria',]$Abundance_r)
```

# Microbiome composition: order level

Exploration of orders in most abundant groups
```{r}
sort(unique(Tax_Data1[Tax_Data1$Phylum=="p__Pseudomonadota",]$Order))
sort(unique(Tax_Data1[Tax_Data1$Phylum=="p__Pseudomonadota",]$Class))

sort(unique(Tax_Data1[Tax_Data1$Phylum=="p__Bacteroidota",]$Order))
sort(unique(Tax_Data1[Tax_Data1$Phylum=="p__Bacteroidota",]$Class))
```

Adjusting the dataset to order level
Illumina
```{r}
Tax_Data1[Tax_Data1$Order=="o__Incertae_Sedis",]$Order<-
				c('c__Gracilibacteria_Inc_Sed','p__NB1_j_Inc_Sed')

Tax_Order1<- aggregate(Tax_Data1[,2:21],by=list(Order=Tax_Data1$Order), sum)
Tax_Order1$Order<-substr(Tax_Order1$Order,4,40)

Tax_Order1<-data.frame(t(Tax_Order1))
colnames(Tax_Order1)<-Tax_Order1[1,]
Tax_Order1<-Tax_Order1[2:21,]
Tax_Order1$month<-substr(rownames(Tax_Order1),3,10)
Tax_Order1$Sample_ID<-rownames(Tax_Order1)
Tax_Order1$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Ord1<- gather(Tax_Order1,Order,Abundance,
                  Gracilibacteria_Inc_Sed:NB1_j_Inc_Sed, factor_key=TRUE)
Tax_Ord1$Abundance_r<-round(as.numeric(Tax_Ord1$Abundance),4)
```

Nanopore
```{r}
Tax_Data2[Tax_Data2$Order=="o__Incertae_Sedis",]$Order<-c('c__OM190_inc._sed.',
			'c__Gracilibacteria_inc._sed.','p__Marinimicrobia_inc._sed.')

Tax_Order2<- aggregate(Tax_Data2[,2:21],by=list(Order=Tax_Data2$Order), sum)
Tax_Order2$Order<-substr(Tax_Order2$Order,4,40)

Tax_Order2<-data.frame(t(Tax_Order2))
colnames(Tax_Order2)<-Tax_Order2[1,]
Tax_Order2<-Tax_Order2[2:21,]
Tax_Order2$month<-substr(rownames(Tax_Order2),3,10)
Tax_Order2$Sample_ID<-rownames(Tax_Order2)
Tax_Order2$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Ord2<- gather(Tax_Order2,Order,Abundance,
                  Gracilibacteria_inc._sed.:Marinimicrobia_inc._sed., factor_key=TRUE)
Tax_Ord2$Abundance_r<-round(as.numeric(Tax_Ord2$Abundance),4)
```

Filtering only 10 orders with the highest abundance
```{r}
Part<-names(sort(tapply(as.numeric(Tax_Ord1$Abundance),Tax_Ord1$Order,mean)))[28:36]
Tax_Ord1$Order_N<-as.character(Tax_Ord1$Order)
Tax_Ord1[!(Tax_Ord1$Order_N %in% Part),]$Order_N<-'Others'
Tax_Ord1$platform<-'Illumina'

Part<-names(sort(tapply(as.numeric(Tax_Ord2$Abundance),Tax_Ord2$Order,mean)))[33:41]
Tax_Ord2$Order_N<-as.character(Tax_Ord2$Order)
Tax_Ord2[!(Tax_Ord2$Order_N %in% Part),]$Order_N<-'Others'
Tax_Ord2$platform<-'Nanopore'
```

Combining datasets
```{r}
Tax_Ord<-rbind(Tax_Ord1,Tax_Ord2)

Tax_Ord_New<-aggregate(Tax_Ord$Abundance_r,list(month=Tax_Ord$month,
			Order=Tax_Ord$Order_N,platform=Tax_Ord$platform),mean)

Tax_Ord_New[Tax_Ord_New$Order=='Gracilibacteria_inc._sed.',]$Order<-
  'Gracilibacteria inc. sed.'

Tax_Ord_New$month<-factor(Tax_Ord_New$month,levels=c('February','March','April','May'))
Tax_Ord_New$Order<-factor(Tax_Ord_New$Order,
	levels = c('Others','Gracilibacteria inc. sed.','Hyphomicrobiales',
		'Chitinophagales','Campylobacterales','Beggiatoales',
		'Verrucomicrobiales','Enterobacterales','Ectothiorhodospirales',
		'Rhodobacterales','Flavobacteriales','Cardiobacteriales'))

sort(tapply(Tax_Ord_New$x,Tax_Ord_New$Order,mean))
```

Figure of bacterial compositon at order level
```{r}
P_Ord<-ggplot(Tax_Ord_New,aes(x = month, y = x, fill = Order)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = c("#DDCC77","#CC6677","#117733","#332288",
		'#993B3B',"#888888","#88CCEE","#AA4499","#44AA99","#999933",
		"#882255","#6699CC")) +
  labs(title= 'B)', x= 'Month')+
	ylab('')+
  theme_minimal()+ 
	theme(axis.text.x = element_text(vjust = 0.5, hjust=1))+
	facet_wrap(~platform)+ 
	theme(text = element_text(size = 17),
	axis.text.x = element_text(angle = 90, hjust = 1))
```

Combined figure

```{r fig6, fig.cap="Microbiome composition: order level"}
P_Ord
```

Exploration of mean abundance for individual groups
```{r}
tapply(Tax_Ord1$Abundance_r,Tax_Ord1$Order,mean)
tapply(Tax_Ord2$Abundance_r,Tax_Ord2$Order,mean)
```

# Adjustment of dataset to genus level

Illumina
```{r}
Tax_Data1$Genus_N<-paste0(Tax_Data1$Genus,'_',Tax_Data1$SampleID)

Tax_Data1[Tax_Data1$Genus_N=='g__Incertae_Sedis_sp1',]$Genus_N<-
  'g__Cardiobacteriaceae inc. sed._sp1'
Tax_Data1[Tax_Data1$Genus_N=="g__NS10_marine_group_sp6",]$Genus_N<-
  'g__Cryomorphaceae NS10 marine group_sp1'
Tax_Data1[Tax_Data1$Genus_N=="g__Incertae_Sedis_sp3",]$Genus_N<-
  'g__Ectothiorhodospiraceae inc. sed._sp1'
Tax_Data1[Tax_Data1$Genus_N=="g__Incertae_Sedis_sp11",]$Genus_N<-
  'g__Flavobacteriaceae inc. sed._sp1'
Tax_Data1[Tax_Data1$Genus_N=="g__Incertae_Sedis_sp8",]$Genus_N<-
  'g__Flavobacteriales NS9 inc. sed._sp1'
Tax_Data1[Tax_Data1$Genus_N=="g__Incertae_Sedis_sp10",]$Genus_N<-
  'g__Rhizobiaceae inc. sed._sp1'

Tax_Genus1<- aggregate(Tax_Data1[,2:21],by=list(Genus=Tax_Data1$Genus_N), sum)
Tax_Genus1$Genus<-substr(Tax_Genus1$Genus,4,40)
```

Nanopore
```{r}
Tax_Data2$Genus_N<-paste0(Tax_Data2$Genus,'_',Tax_Data2$SampleID)

Tax_Data2[Tax_Data2$Genus_N=='g__Incertae_Sedis_sp1',]$Genus_N<-
  'g__Cardiobacteriaceae inc. sed._sp1'
Tax_Data2[Tax_Data2$Genus_N=="g__NS10_marine_group_sp6",]$Genus_N<-
  'g__Cryomorphaceae NS10 marine group_sp1'
Tax_Data2[Tax_Data2$Genus_N=="g__Incertae_Sedis_sp3",]$Genus_N<-
  'g__Ectothiorhodospiraceae inc. sed._sp1'
Tax_Data2[Tax_Data2$Genus_N=="g__Incertae_Sedis_sp11",]$Genus_N<-
  'g__Flavobacteriaceae inc. sed._sp1'
Tax_Data2[Tax_Data2$Genus_N=="g__Incertae_Sedis_sp8",]$Genus_N<-
  'g__Flavobacteriales NS9 inc. sed._sp1'
Tax_Data2[Tax_Data2$Genus_N=="g__Incertae_Sedis_sp10",]$Genus_N<-
  'g__Rhizobiaceae inc. sed._sp1'

Tax_Genus2<- aggregate(Tax_Data2[,2:21],by=list(Genus=Tax_Data2$Genus_N), sum)
Tax_Genus2$Genus<-substr(Tax_Genus2$Genus,4,40)
```

Tidying the data set
Illumina
```{r}
Tax_Genus1<-data.frame(t(Tax_Genus1))
colnames(Tax_Genus1)<-Tax_Genus1[1,]
Tax_Genus1<-Tax_Genus1[2:21,]
Tax_Genus1$month<-substr(rownames(Tax_Genus1),3,10)
Tax_Genus1$Sample_ID<-rownames(Tax_Genus1)
Tax_Genus1$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Gen1<- gather(Tax_Genus1,Genus,Abundance,Acinetobacter_sp196:Zobellia_sp171, factor_key=TRUE)
Tax_Gen1$Abundance_r<-round(as.numeric(Tax_Gen1$Abundance),4)
```

Nanopore
```{r}
Tax_Genus2<-data.frame(t(Tax_Genus2))
colnames(Tax_Genus2)<-Tax_Genus2[1,]
Tax_Genus2<-Tax_Genus2[2:21,]
Tax_Genus2$month<-substr(rownames(Tax_Genus2),3,10)
Tax_Genus2$Sample_ID<-rownames(Tax_Genus2)
Tax_Genus2$Sample_ID<-paste0(rep(1:5,4),'_',rep(c('Feb','Mar','Apr','May'),each=5))

Tax_Gen2<- gather(Tax_Genus2,Genus,Abundance,Acinetobacter_sp196:Yoonia_sp23, factor_key=TRUE)
Tax_Gen2$Abundance_r<-round(as.numeric(Tax_Gen2$Abundance),2)
```

Filtering only 10 genus with the highest abundance
```{r}
Part<-names(sort(tapply(as.numeric(Tax_Gen1$Abundance),Tax_Gen1$Genus,mean)))[124:132]
Tax_Gen1$Genus_N<-as.character(Tax_Gen1$Genus)
Tax_Gen1[!(Tax_Gen1$Genus_N %in% Part),]$Genus_N<-'Others'
Tax_Gen1$platform<-'Illumina'

Part<-names(sort(tapply(as.numeric(Tax_Gen2$Abundance),Tax_Gen2$Genus,mean)))[126:134]
Tax_Gen2$Genus_N<-as.character(Tax_Gen2$Genus)
Tax_Gen2[!(Tax_Gen2$Genus_N %in% Part),]$Genus_N<-'Others'
Tax_Gen2$platform<-'Nanopore'
```

Combining datasets
```{r}
Tax_Gen<-rbind(Tax_Gen1,Tax_Gen2)

Tax_Gen_New<-aggregate(Tax_Gen$Abundance_r,list(month=Tax_Gen$month,
			Genus=Tax_Gen$Genus_N,platform=Tax_Gen$platform),mean)
```

Tidying the dataset
```{r}
Tax_Gen_New1<-aggregate(Tax_Gen$Abundance_r,list(month=Tax_Gen$month,Sample=Tax_Gen$Sample_ID,
			Genus=Tax_Gen$Genus_N,platform=Tax_Gen$platform),mean)

New_Illum<-Tax_Gen_New1[Tax_Gen_New1$platform=='Illumina',]
New_Nan<-Tax_Gen_New1[Tax_Gen_New1$platform=='Nanopore',]
```

# Venn diagrams: platforms

```{r}
Tax_Phylum1<- aggregate(Tax_Data1[,2:21],by=list(Phylum=Tax_Data1$Phylum), sum)
Tax_Genus1<- aggregate(Tax_Data1[,2:21],by=list(Genus=Tax_Data1$Genus_N), sum)
Ill_Phyl<-substr(Tax_Phylum1$Phylum,4,30)
Ill_Gen<-substr(Tax_Genus1$Genus,4,30)

Tax_Phylum2<- aggregate(Tax_Data2[,2:21],by=list(Phylum=Tax_Data2$Phylum), sum)
Tax_Genus2<- aggregate(Tax_Data2[,2:21],by=list(Genus=Tax_Data2$Genus_N), sum)
Nan_Phyl<-substr(Tax_Phylum2$Phylum,4,20)
Nan_Gen<-substr(Tax_Genus2$Genus,4,30)
```

Calculate abundance
```{r}
Tax_Gen_Il<-aggregate(as.numeric(Tax_Gen1$Abundance),list(Genus=Tax_Gen1$Genus),mean)
colnames(Tax_Gen_Il)[2]<-'Illumina'
Tax_Gen_Na<-aggregate(as.numeric(Tax_Gen2$Abundance),list(Genus=Tax_Gen2$Genus),mean)
colnames(Tax_Gen_Na)[2]<-'Nanopore'
```

Merging of datasets
```{r}
Tax_Gen_O<-merge(Tax_Gen_Il,Tax_Gen_Na,by='Genus',all=TRUE)
Tax_Gen_O[is.na(Tax_Gen_O)]<-0
Tax_Gen_O$Overall_Abun<-rowMeans(Tax_Gen_O[,2:3])
```

Calculation of abundance for shared genera
```{r}
x <- list(Illumina = Ill_Gen, Nanopore = Nan_Gen)
ItemsList <- gplots::venn(x, show.plot = FALSE)

Abun_list<-attributes(ItemsList)$intersections$`Illumina:Nanopore`
sum(Tax_Gen_O[Tax_Gen_O$Genus %in% Abun_list,]$Overall_Abun)

Tax_Gen_O$Genus_N<-Tax_Gen_O$Genus
Tax_Gen_O$Genus<-sub("_s[^_]+$", "", as.vector(Tax_Gen_O$Genus))
Tax_Gen_O$SampleID<-paste0('s',sub(".*_s", "", as.vector(Tax_Gen_O$Genus_N)))
```

Figure of venn diaggram for platforms
```{r}
Venn_plat<-ggvenn(x,fill_color = c("#D55E00","#0072B2"),stroke_size = 0.5,
                  fill_alpha = 0.8,text_size = 5.5,
	show_percentage=FALSE,set_name_size = 0)+
	ggplot2::annotate(geom="text",x=0.06, y=-0.2,label="98.6 %",size = 5)+
	ggplot2::theme(text = element_text(size = 8),
		legend.position = "none",
		plot.title = element_blank())
```

```{r fig7, fig.cap="Venn diagram: platforms"}
Venn_plat <- Venn_plat + theme(plot.tag = element_text(size = 16, face = "plain"))

Venn_plat  +
  plot_annotation(tag_levels = list(c("A)")))
```

# Venn diagram: Comparison of months

```{r}
Ill_Feb<-Tax_Genus1[,c(1:6)][apply(Tax_Genus1[,c(2:6)],1,sum)>0,]
Ill_Mar<-Tax_Genus1[,c(1,7:11)][apply(Tax_Genus1[,c(7:11)],1,sum)>0,]
Ill_Apr<-Tax_Genus1[,c(1,12:16)][apply(Tax_Genus1[,c(12:16)],1,sum)>0,]
Ill_May<-Tax_Genus1[,c(1,17:21)][apply(Tax_Genus1[,c(17:21)],1,sum)>0,]

Nan_Feb<-Tax_Genus2[,c(1:6)][apply(Tax_Genus2[,c(2:6)],1,sum)>0,]
Nan_Mar<-Tax_Genus2[,c(1,7:11)][apply(Tax_Genus2[,c(7:11)],1,sum)>0,]
Nan_Apr<-Tax_Genus2[,c(1,12:16)][apply(Tax_Genus2[,c(12:16)],1,sum)>0,]
Nan_May<-Tax_Genus2[,c(1,17:21)][apply(Tax_Genus2[,c(17:21)],1,sum)>0,]
```

Calculation of abundance
```{r}
Tax_Gen_I<-aggregate(as.numeric(Tax_Gen1$Abundance),list(Genus=Tax_Gen1$Genus),mean)
colnames(Tax_Gen_I)[2]<-'Illumina'
Tax_Gen_N<-aggregate(as.numeric(Tax_Gen2$Abundance),list(Genus=Tax_Gen2$Genus),mean)
colnames(Tax_Gen_N)[2]<-'Nanopore'
```

Illumina
```{r}
Illum <- list(February=Ill_Feb$Genus,March=Ill_Mar$Genus,April=Ill_Apr$Genus,
		May=Ill_May$Genus)

ItemsList <- gplots::venn(Illum, show.plot = FALSE)
Abun_list<-attributes(ItemsList)$intersections$`February:March:April:May`
Abun_list<-substr(Abun_list,4,30)
```

Nanopore
```{r}
Nanop <- list(February=Nan_Feb$Genus,March=Nan_Mar$Genus,April=Nan_Apr$Genus,
		May=Nan_May$Genus)

ItemsList <- gplots::venn(Nanop, show.plot = FALSE)
Abun_list<-attributes(ItemsList)$intersections$`February:March:April:May`
Abun_list<-substr(Abun_list,4,30)
```

Figures of venn diagram for each platform separately
```{r}
P1<-ggvenn(Illum, 
  fill_color = c("#CC6677","#DDCC77","#117733","#332288"),
  stroke_size = 0.5,set_name_size = 2.8,fill_alpha = 0.5,
	show_percentage=FALSE)+
	ggplot2::ggtitle('A)  Illumina') +
	ggplot2::annotate(geom="text",x=0.06, y=-0.85,label="99.3 %")

P2<-ggvenn(Nanop, 
  fill_color = c("#CC6677","#DDCC77","#117733","#332288"),
  stroke_size = 0.5,set_name_size = 2.8,fill_alpha = 0.5,
	show_percentage=FALSE)+
	ggplot2::ggtitle('B)  Nanopore') +
	ggplot2::annotate(geom="text",x=0.06, y=-0.85,label="99.5 %")
```

Both figures
```{r fig8, fig.cap="Venn diagram: months"}
(P1 | P2 )
```

# LDA and LefSe analysis

Adjustment of data for analysis
```{r}
colnames(TAX1)[c(1)]<-c('Kingdom')
Ill_phylo_filt_1 <- phyloseq(otu_table(Ill_atu_filt), TAX1, SAM1)

colnames(TAX2)[c(1)]<-c('Kingdom')
Nan_phylo_filt_1 <- phyloseq(otu_table(Nan_atu_filt), TAX2, SAM2)
```

LefSe analysis
```{r}
result_Illum <- run_lefse(Ill_phylo_filt_1,group='time', taxa_rank='Genus')
result_Nano <- run_lefse(Nan_phylo_filt_1,group='time', taxa_rank='Genus')
```

Compile results into a single dataset
```{r}
LDA_Illu<-data.frame(Group=marker_table(result_Illum)$feature,
                     Month=marker_table(result_Illum)$enrich_group,
	LDA=marker_table(result_Illum)$ef_lda,Platform='Illumina')

LDA_Nan<-data.frame(Group=marker_table(result_Nano)$feature,
                    Month=marker_table(result_Nano)$enrich_group,
	LDA=marker_table(result_Nano)$ef_lda,Platform='Nanopore')

LDA<-rbind(LDA_Nan,LDA_Illu)
```

Filtering of groups with LDA scores higher than 4
```{r}
LDA_filt <- LDA[ave(LDA$LDA >= 4, LDA$Group, FUN = any), ]

LDA_filt$Group<-substr(LDA_filt$Group,4,50)
LDA_filt$Group<-factor(LDA_filt$Group,levels=c(sort(unique(LDA_filt$Group))))
```

Adjustment of dataset for plotting
```{r}
levels(LDA_filt$Group)[c(2:7,10:12)]<-c("Cardiobacteriaceae inc. sed.",
	"Cyanobacteriia chloroplast inc. sed.","Cryomorphaceae inc. sed.",
	"Ectothiorhodospiraceae inc. sed.","Flavobacteriaceae inc. sed.",
	"Gracilibacteria inc. sed.","Cryomorphaceae NS10 marine group",
	"Flavobacteriales NS9 inc. sed.","Rhizobiaceae inc. sed."
	)
	
LDA_filt_compl <- LDA_filt %>%
  complete(Group, Platform, fill = list(LDA = 0))
```

Figure of LefSe analysis
```{r}
LefSe<-ggplot(LDA_filt_compl, aes(LDA, Group, fill = Platform)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.85), width = 0.85) +
  scale_fill_manual(values = c("#D55E00", "#0072B2")) +
  labs(y = '', x = 'LDA score [log10]') +
  theme_minimal() +
  theme(text = element_text(size = 15)) +
  geom_vline(xintercept = 4, linetype = "dashed") +
  scale_y_discrete(labels = c(
	expression(italic("Arcobacter")),
	expression(italic("Cardiobacteriaceae inc. sed.")),
      expression("Cyanobacteriia chloroplast " * italic("inc. sed.")),
	expression(italic("Cryomorphaceae inc. sed.")),
	expression(italic("Ectothiorhodospiraceae inc. sed.")),
	expression(italic("Flavobacteriaceae inc. sed.")),
	expression(italic("Gracilibacteria inc. sed.")),
	expression(italic("Halocynthiibacter")),
	expression(italic("Leucothrix")),
      expression(italic("Cryomorphaceae") * " NS10 marine group"),
      expression("Flavobacteriales NS9 " * italic("inc. sed.")),
	expression(italic("Rhizobiaceae inc. sed.")),
	expression(italic("Rubritalea")),
	expression(italic("Ruegeria")),
	expression(italic("Salinimicrobium")),
	expression(italic("Tenacibaculum")),
	expression(italic("Vibrio"))))
```

The final figure together with venn diagram of platforms
```{r fig9, fig.cap="LefSe analysis"}
LefSe    <- LefSe + theme(plot.tag = element_text(size = 16, face = "plain"))

(LefSe) + 
  plot_annotation(tag_levels = list(c("B)")))
```

# Identification of pathogen-associated genera

```{r}
Detected<-sort(substr(unique(observation$Genus),4,25))
```

List of pathogen-associated genera
```{r}
Pathogens<-c('Acinetobacter','Aerococcus','Aeromonas','Aliivibrio','Aquaspirillum',
		'Arcobacter','Carnobacterium','Chlamydia','Chryseobacterium',
		'Citrobacter','Clostridium','Delftia','Edwardsiella','Enterobacter',
		'Erysipelothrix','Escherichia','Flavobacterium','Flectobacillus',
		'Francisella','Hafnia','Hahella','Halomonas','Janthinobacterium',
		'Klebsiella','Lactococcus','Moritella','Mycobacterium','Mycoplasma',
		'Nocardia','Pantoea','Pasteurella','Photobacterium','Piscirickettsia',
		'Plesiomonas','Providencia','Pseudoalteromonas','Pseudomonas',
		'Renibacterium','Salmonella','Serratia','Shewanella','Stenotrophomonas',
		'Streptococcus','Tenacibaculum','Vagococcus','Vibrio','Weissella','Yersinia')
```

Matching detected genera to the list of fish pathogens
```{r}
pattern<- c(Pathogens[1]) 
myStrings <- observation$Genus
Det_Pathogens<-data.frame()

Det_Pathogens<-observation[observation$Genus %in% c(myStrings[grepl(pattern, myStrings)]),]

for (i in 2:48){
	pattern<- c(Pathogens[i]) 
	myStrings <- observation$Genus
	if(length(myStrings[grepl(pattern, myStrings)])>0){
	Part<-observation[observation$Genus %in% c(myStrings[grepl(pattern, myStrings)]),]
	Det_Pathogens<-rbind(Det_Pathogens,Part)
	}}

Det_Pathogens$Genus_N<-substr(Det_Pathogens$Genus,4,20)
```

Matching the list of detected pathogens to abundance data at genus level
```{r}
Tax_Gen_New1<-aggregate(Tax_Gen$Abundance_r,list(Month=Tax_Gen$month,
			Genus=Tax_Gen$Genus,platform=Tax_Gen$platform),mean)
colnames(Tax_Gen_New1)[4]<-'mean'
Tax_Gen_New2<-aggregate(Tax_Gen$Abundance_r,list(Month=Tax_Gen$month,
			Genus=Tax_Gen$Genus,platform=Tax_Gen$platform),function(x) {sd(x)/sqrt(length(x))})
colnames(Tax_Gen_New2)[4]<-'se'
Tax_Gen_New<-merge(Tax_Gen_New1,Tax_Gen_New2,by=c('Month','Genus','platform'),all=TRUE)

Tax_Gen_New$Genus_N<-sub("_s[^_]+$", "", as.vector(Tax_Gen_New$Genus))
Tax_Gen_Pat<- Tax_Gen_New[Tax_Gen_New$Genus_N %in% c(Det_Pathogens$Genus_N),]

sort(tapply(Tax_Gen_Pat$mean,Tax_Gen_Pat$Genus_N,sum))
Tax_Gen_Pat1<-Tax_Gen_Pat[Tax_Gen_Pat$Genus_N %in% c('Tenacibaculum','Vibrio','Arcobacter'),]

Tax_Gen_Pat1$Genus_N<-factor(Tax_Gen_Pat1$Genus_N,
	levels = c('Tenacibaculum','Vibrio','Arcobacter'))
Tax_Gen_Pat1$Month<-factor(Tax_Gen_Pat1$Month,
	levels = c('February','March','April','May'))
```

Testing of differences in abundance between months
```{r}
Abun<-aggregate(Tax_Gen$Abundance_r,list(Month=Tax_Gen$month,sample=Tax_Gen$Sample_ID,
			Genus=Tax_Gen$Genus,platform=Tax_Gen$platform),mean)
Abun$Genus_N<-sub("_s[^_]+$", "", as.vector(Abun$Genus))
Abun_Pat<- Abun[Abun$Genus_N %in% c(Det_Pathogens$Genus_N),]
Abun_Pat1<-Abun_Pat[Abun_Pat$Genus_N %in% c('Tenacibaculum','Vibrio','Arcobacter'),]

Abun_Pat1$Devide<-paste(Abun_Pat1$Genus_N,Abun_Pat1$platform)

for(i in 1:length(unique(Abun_Pat1$Devide))){
	A<-Abun_Pat1[Abun_Pat1$Devide %in% unique(Abun_Pat1$Devide)[i],]
	Krus<-kruskal.test(x ~ Month, data = A)
	if(Krus$p.value<0.05){
		print(unique(Abun_Pat1$Devide)[i])
		print(pairwise.wilcox.test(A$x,A$Month,p.adjust.method = "BH"))
			}else{
		print('NO')}}

Ill_Arc<-c('a','a','a','a')
Ill_Ten<-c('a','b','ab','a')
Ill_Vib<-c('ab','a','ab','b')
Nan_Arc<-c('ab','a','b','a')
Nan_Ten<-c('a','a','a','b')
Nan_Vib<-c('a','a','a','a')
```

Preparation of data for plotting
```{r}
summary_data <- Tax_Gen_Pat1 %>%
  group_by(platform,Genus_N,Month) %>%
  summarise(mean = mean(mean, na.rm = TRUE),
            se = se,.groups = 'drop')

summary_data <- summary_data %>%
	mutate(
		Month = factor(Month, levels = c("February", "March", "April", "May")),
		Genus_N = factor(Genus_N, levels = sort(unique(Genus_N))),
		platform = factor(platform, levels = c("Illumina", "Nanopore")) ) %>%
			arrange(platform, Genus_N, Month)

label_data <- summary_data %>%
  mutate(label = c(Ill_Ten,Ill_Vib,Ill_Arc,Nan_Ten,Nan_Vib,Nan_Arc),  
         y = mean + se + 0.15) 
```

```{r}
Abun_Pat1$Genus_N<-factor(Abun_Pat1$Genus_N,levels=levels(Tax_Gen_Pat1$Genus_N))
Abun_Pat1$Month<-factor(Abun_Pat1$Month,levels=c('February','March','April','May'))

Tax_Gen_Pat1$month<-factor(Tax_Gen_Pat1$Month,levels=c("February", "March", "April", "May"))
label_data<-as.data.frame(label_data)
label_data$Month<-factor(label_data$Month,levels=c("February", "March", "April", "May"))

Illumina<-Tax_Gen_Pat1[Tax_Gen_Pat1$platform=='Illumina',]
label_d<-label_data[label_data$platform=='Illumina',]
Illumina$Genus_N<-factor(Illumina$Genus_N,levels=c('Arcobacter','Tenacibaculum','Vibrio'))
label_data$Genus_N<-factor(Illumina$Genus_N,levels=c('Arcobacter','Tenacibaculum','Vibrio'))
```

Figure of pathogenic-genera abundance
```{r fig10, fig.cap="Abundance of pathogenic genera"}
ggplot(Illumina,aes(x =Genus_N, y = mean, color = Month)) +
	geom_point(position=position_dodge(width = .5),size=4)  +
	labs(x= 'Genus',y= 'Proportion of the total abundance')+
	theme_minimal()+
	geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.4,
                 position=position_dodge(.5)) + 
	scale_color_manual(values = c("#CC6677","#DDCC77","#88CCEE","#117733")) +
#	geom_point(data = Abun_Pat1, 
#            mapping = aes(x = Genus_N, y = x, color = Month),
#		position=position_dodge(width = .5),size=1)+
	geom_text(data = label_d,
            aes(x = Genus_N, y = mean+se+0.5, label = label,color=Month,group=Month),
            position = position_dodge(width = 0.5),
            inherit.aes = FALSE,
            size = 5)+
	theme(text = element_text(size = 17),
		axis.text.x = element_text(face = "italic"))
```

# List of detected pathogen-associated genera

Using the original, unfiltered dataset, the pathogen-associated genera were identified
```{r}
Sequen$Sample_ID<-rownames(Sequen)
Count_Path<-Sequen[Sequen$Sample_ID %in% c(Det_Pathogens$SampleID),]

Count_Path<-data.frame(t(Count_Path[,1:40]))
Count_Path$Month<-c(rep('Feb',5),rep('Mar',5),rep('Apr',5),rep('May',5))
Count_Path$Platform<-c(rep('Illumina',20),rep('Nanopore',20))
Count_Path<- aggregate(Count_Path[,1:32],by=list(Platform=Count_Path$Platform), sum)

Tax<-observation[observation$SampleID %in% colnames(Count_Path),]
colnames(Count_Path)[2:33]<-Tax$Genus

Count_Path_F<- gather(Count_Path,Genus,Count,g__Arcobacter:g__Plesiomonas, factor_key=TRUE)
Count_Path_F$Genus<-substr(Count_Path_F$Genus,4,20)
```

Filtering for genera with at least 10 reads in all samples
```{r}
genus_platform <- with(Count_Path_F, tapply(Count, list(Genus, Platform), sum))
keep_genera <- rownames(genus_platform)[apply(genus_platform >= 10, 1, any)]
Count_Path_F3<-Count_Path_F[Count_Path_F$Genus %in% keep_genera, ]

Count_Path_F4<- spread(Count_Path_F3,Platform,Count)
Count_Path_F4
```

Number of detected genera
```{r}
nrow(Count_Path_F4)
nrow(Count_Path_F4[Count_Path_F4$Illumina>10,])
nrow(Count_Path_F4[Count_Path_F4$Nanopore>10,])
```

Calculation of overall abundance for pathogen-associated genera
```{r}
Selec<-Tax_Data[Tax_Data$SampleID %in% Tax$SampleID,]
Selec_Ill<-Selec[,c(2:21)]
rownames(Selec_Ill)<-Selec[,c(47)]
Selec_Nan<-Selec[,c(22:41)]
rownames(Selec_Nan)<-Selec[,c(47)]

apply(Selec_Ill,1,mean)
apply(Selec_Nan,1,mean)
```

# Species level identification of pathogens

Loading data: Microbiome pathogens
```{r}
Pathog<-read.table("Sealice_Microbiome_Pathogens.csv",
	header = TRUE,sep = ';',fill = TRUE,dec = ",",na.strings = "NA")
```

Calculation of abundance for individual pathogens and months

```{r}
Pathog<-Pathog[Pathog$Species %in% c('Acinetobacter johnsonii','Aeromonas sobria',
		'Aliivibrio logei','Aliivibrio wodanis','Moritella marina',
		'Pseudomonas fluorescens','Pseudomonas koreensis',
		'Tenacibaculum dicentrarchi','Tenacibaculum maritimum',
		'Tenacibaculum ovolyticum','Tenacibaculum soleae',
		'Vibrio alginolyticus','Vibrio anguillarum','Vibrio splendidus',
		'Vibrio tapetis'),]


Patogen<-data.frame(Species=Pathog$Species,Ab_Total=apply
                    (Pathog[,c(3:22)],1,mean),
	Ab_Feb=apply(Pathog[,c(3:7)],1,mean),Ab_Mar=apply
	(Pathog[,c(8:12)],1,mean),
	Ab_Apr=apply(Pathog[,c(13:17)],1,mean),Ab_May=apply
	(Pathog[,c(18:22)],1,mean))

Patogen[Patogen$Ab_Feb==0,]$Ab_Feb<-NA
Patogen[Patogen$Ab_Mar==0,]$Ab_Mar<-NA
Patogen[Patogen$Ab_Apr==0,]$Ab_Apr<-NA
Patogen[Patogen$Ab_May==0,]$Ab_May<-NA
```

```{r table1, results="asis", tab.cap="Abundance of putative pathogens"}
knitr::kable(Patogen)
```


